Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2018-11-05 article_springer/article_springer.Rmd
Section 4

: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999.

@@

In the classical POT approach a GPD distribution is fitted to the data. Due to the Theorem in the last Section we know that for high threshholds the waiting times between the exceedances are Mittag-Leffler distributed in case of heavy tailed waiting times. Hence, we discuss in this section first parameter estimation for Mittag-Leffler distributions.

The first method proposed for parameter estimation of the Mittag-Leffler distribution was the fractional moment estimator by [@kozubowski2001]. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. [@Cahoy2010] proposed an estimator based on the log-Moments which doesn't need prior knowledge. Furthermore, they showed in a simulation study that the log-Moments estimator outclasses the fractional moment estimator regarding bias and root mean squared error (RMSE). Due to the form of the density function there exists no closed form for the Maximum Likelihood estimator (MLE). In the RPackage `MittagLeffleR` Maximum Likelihood estimation is implemented via numerical optimization. The MLE outperforms the log-moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computational intensive and the differences in bias and RMSE between the two methods are very small. In Figure \ref{fig:MSE} one can see that both methods work quite well even for very low sample sizes.    
Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators such as the Hill estimator. However, these methods are less statistically efficient since they only use a portion of the information contained in the data, as was already mentionded by [@kozubowski2001]. Addtionally they need a specification of a tuning parameter. In XX one can see a Hill plot for differet k's.

@@

```{r LRT power, echo=FALSE, fig.cap="Type One error for bootstrapped LRT for different sample sizes", out.width = '100%'}
knitr::include_graphics("Plots_LRT_boot.pdf")
```

```{r LRT type one error, echo=FALSE, fig.cap="Type One error for bootstrapped LRT for different sample sizes", out.width = '100%'}
knitr::include_graphics("Plots_LRT_boot_typeOneError.pdf")
```

>>>>

: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999. In @Anderson1987  a similar result is shown with a different choice of scaling constant.

@@

The classical POT approach is based on the fact, that exceedances above a high threshhold are assymptotically GPD distributed, hence a GPD distribution is fitted to the exceedances above a high enogh threshhold. Due to the Theorem in the last Section we know that for high threshholds the times between the exceedances are assymptotically Mittag-Leffler distributed in case of heavy tailed waiting times between the observations. Hence, before we talk about inference for the exceedances in the next section, we want to have a look at inference for Mittag-Leffler distributions in the folowing.

The first method proposed for the estimation of the Mittag-Leffler distribution parameters was the fractional moment estimator by [@kozubowski2001]. Unlike the first moments, the fractional moments of order $p$ for $p<\beta$ exist and can get calculated. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. [@Cahoy2010] proposed an estimator based on the log-Moments which doesn't need prior knowledge. Furthermore, they showed in a simulation study that the log-Moment outclasses the fractional moment estimator regarding bias and root mean squared error (RMSE).  
Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the RPackage `MittagLeffleR` Maximum Likelihood estimation is implemented via numerical optimization. The MLE outperforms the log-moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computational intensive and the differences in bias and RMSE between the two methods are very small. In Figure \ref{fig:MSE} one can see that both methods work quite well even for very low sample sizes.    
Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator, which is defined by 
\begin{align*}
H_{k,n}= \frac{1}{k} \sum_{i=0}^{k-1} \log \frac{X_{n-i,n}}{X_{n-k,n}}. 
\end{align*}
However, these methods are less statistically efficient since they only use a portion of the information contained in the data, as was already mentioned by [@kozubowski2001]. Addtionally, they need a specification of a tuning parameter. In Figure one can see a Hill plot for different $k's$.

@@

```{r LRT power, echo=FALSE, fig.cap="Power for bootstrapped LRT for different sample sizes", out.width = '100%'}
knitr::include_graphics("Plots_LRT_boot.pdf")
```
